\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
Wrap up of my first pass.  A few diag rebuilds.

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang--Mills, that
  unifies many of the color-dual theories studied in the
  literature, and provides a prescriptive approach for constructing color-dual numerators at
  one-loop directly from Feynman rules. Consideration of this
  theory highlights that the lifting from one-loop color-kinematics
  dual numerators to two loops is just as challenging as the progression
  from tree-level to one-loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  nonlinear sigma model (NLSM) numerators using a $D$-dimensional color-dual bootstrap.  The double-copy of the constructed NLSM numerators with
  themselves, pure Yang--Mills, and $\mathcal{N}=4$
  super-Yang-Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our constructive bootstrap, we also find that the most general \emph{local} two-loop four-point numerator for pure Yang-Mills fails to satisfy global color-kinematics duality. We pinpoint the failure to the bow-tie unitarity cut, and discuss a path towards \emph{non-local} construction of color-dual integrands at generic loop order.

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
  numerator
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM, $\phi^3$' which was redundant.  I don't mind if we say it here because I want people to tell people that these are the kinds of theories we're thinking about.  We could omit it but then we'd have to say that the theory has at least one color factor and is probably an adjoint type theory and at that point I'd rather just give an example.} -- can be written as
\begin{equation}
\mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

Once color-dual numerators $n_\Gamma$ have been found, they can be
used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
the double-copy theory \be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)}
~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma}.
\end{equation}
Importantly, the $\tilde{n}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the KLT relations but there is significant evidence that
the double copy persists to all loop orders \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv, Carrasco:2021otn, KLT, KiermaierTalk}.
In addition to pointing at deep
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_m^L$ than traditional methods like
Feynman rules.

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, BernHuangReview, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022jln}.  \ace{Here is a semi-random
  collection of unitarity-related citations.  Probably want to bulk it
  up a bit more.}  Recently, generalized unitarity was optimized for
even-multiplicity theories like NLSM \cite{Carrasco:2023qgz}.

@@

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate that many separate
color-orderings of the same cut. \jm{Don't you need $(n-2)!$ for each
  blob in the cut?  So you need to evaluate $(n_1-2)! (n_2-2)!...$
  cuts where $n_1$, $n_2$... are the multiplicities of the
  blobs.}\ace{Yes.  Lets figure out a way to word it.} As an example,
consider the ``bowtie'' cut topology -- one of the factorization
channels of the iterated unitarity cut -- which has two topologically
distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.  \ace{One of the
  diagrams here got shifted around somehow.  Probably need to rebuild
  the whole fig manifest}

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the
four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional
two-form studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the
NMHV Yang-Mills Lagrangian.  After the introduction of the 2-form, the
numerators now satisfy the Jacobi identity
\begin{equation}
  {}_1\langle
  V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0
\end{equation}
where the new cubic Lagrangian takes the form
\begin{equation}
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .
\end{equation}
Of course, higher multiplicity would require further redefinition of
the three-point vertex to satisfy Jacobi identites on all internal
edges. While introducing successively higher spin states could in
principle work at tree level, we will see that it is not consistent
with what we find at general loop order. We will comment on this in
\cref{sec:Discussion} \ace{Is this the sec you are thinking of?}. For
now, we will study how to construct one-loop integrands consistent
with the color-kinematics by isolating the cubic sector of the state
sum above.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box (2box) and penta-triangle (pt) as the
Jacobi basis graphs and building a the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang--Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form:
\begin{equation}
\mathcal{N}_{\text{2box}},\mathcal{N}_{\text{pt}}\in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  In our labeling of the two loop Jacobi-basis
diagrams, shown in \cref{eq:jacobi-basis}, we find that the power
counting bound proposed by Bern, Davies, and Nohle leads to an initial
ansatz with 10,000 terms in the double-box, and 9849 terms for the
penta-triangle.  Without power counting restrictions, both the
double-box and the penta-triangle ansatz have an ansatz with 10,010
terms each.  Imposing diagram automorphisms and maximal-cut gauge
invariance on the two basis diagrams, we reduce the number of terms to
XX for the double-box, and YY for the penta-triangle.  In
Ref.~\cite{Bern:2015ooa}, the unitarity matching constraints were
imposed using a spanning set of minimal physical cuts, shown in
\cref{fig:ym-spanning}.  Instead of also directly going to the
spanning cuts, we work within the framework of the Method of Maximal
Cuts \citepls{} in order to identify the simplest cut that is in
tension with the kinematic Jacobi relations.

\begin{figure}
  \begin{center}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \LMCut
      \end{center}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \PhysicalCutOne{}{}{}{}
      \end{center}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \PhysicalCutTwo{}{}{}{}
      \end{center}
    \end{subfigure}
  \end{center}
  \caption{The three spanning physical unitarity cuts of pure
    Yang--Mills at 2 loops}
  \label{fig:ym-spanning}
\end{figure}

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang--Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut. \jm{Insert clip of Gordon Ramsay saying ``delicious'' here.}

>>>>

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing color-dual
  numerators at one-loop directly from Feynman rules. Consideration of
  this theory highlights that the lifting from one-loop
  color-kinematics dual numerators to two loops is just as challenging
  as the progression from tree-level to one-loop.  As a first step in
  understanding the two-loop challenge, we construct color-dual
  two-loop four-point nonlinear sigma model (NLSM) numerators using a
  $D$-dimensional color-dual bootstrap.  The double-copy of the
  constructed NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known
  physical unitarity cuts of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These
  NLSM numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our constructive bootstrap \ace{I am
    really hesitant to call the bootstrap constructive.  While it
    technically constructs, ``constructive'' has some additional
    connotations}, we also find that the most general \emph{local}
  two-loop four-point numerator for pure Yang-Mills fails to satisfy
  global color-kinematics duality. We pinpoint the failure to the
  bow-tie unitarity cut, and discuss a path towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
  numerator
\item \ace{I am starting to get a bit concerned about the overuse of the
  phrase ``color-dual''}
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM,
  $\phi^3$' which was redundant.  I don't mind if we say it here
  because I want people to tell people that these are the kinds of
  theories we're thinking about.  We could omit it but then we'd have
  to say that the theory has at least one color factor and is probably
  an adjoint type theory and at that point I'd rather just give an
  example.} -- can be written as
\begin{equation}
  \mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over only purely cubic graphs $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu,
    Carrasco:2023vjg, Carrasco:2020ywq}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

Once color-dual numerators $n_\Gamma$ have been found, they can be
used to replace the color factors ($c_\Gamma \to n_\Gamma$) in a
\emph{different} amplitude to produce the corresponding amplitude in
the double-copy theory \be\label{DCconstruction}
\tilde{\mathcal{A}}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma} ~~ \xrightarrow{(c_\Gamma \to n_\Gamma)}
~~ \mathcal{M}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int
\frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{n_\Gamma
  \tilde{n}_\Gamma}{d_\Gamma}.
\end{equation}
Importantly, the $\tilde{n}_\Gamma$ do not need to respect
color-kinematics duality.  The prototypical example of the double-copy
is that YM double copied with YM results in gravity where the gauge
invariance of each separate gluon produces the diffeomorphism
invariance of the graviton.  The double copy can be proven at tree
level using the KLT relations but there is significant evidence that
the double copy persists to all loop orders \cite{FiveLoopN4,
  GeneralizedDoubleCopyFiveLoops, Bern:2018jmv, Carrasco:2021otn, KLT,
  KiermaierTalk}.  In addition to pointing at potential undiscovered
structure hidden in the Lagrangian for gravity, the double copy is
immensely useful at a utilitarian level since it is much more
efficient at producing $\mathcal{M}_m^L$ than traditional methods like
Feynman rules.

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, BernHuangReview, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}.  \ace{Here is a semi-random
  collection of unitarity-related citations.  Probably want to bulk it
  up a bit more.}  Recently, generalized unitarity was optimized for
even-multiplicity theories like NLSM \cite{Carrasco:2023qgz}.

@@

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- which has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \label{eq:ym-lag}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge
\begin{equation}
T^k_{ij} = (\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic generator is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations
\begin{equation}
V^{(123)}_{\text{YM}} = T^3_{12} +\text{cyc}(123).
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
The remaining terms can be absorbed into the definition of the
four-point kinematic numerator
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
where $V_{\text{YM}+B}$ indicates the introduction of an additional
two-form studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the
NMHV Yang-Mills Lagrangian.  After the introduction of the 2-form, the
numerators now satisfy the Jacobi identity
\begin{equation}
  {}_1\langle
  V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0
\end{equation}
where the new cubic Lagrangian takes the form
\begin{equation}
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .
\end{equation}
Of course, higher multiplicity would require further redefinition of
the three-point vertex to satisfy Jacobi identites on all internal
edges. While introducing successively higher spin states could in
principle work at tree level, we will see that it is not consistent
with what we find at general loop order. We will comment on this in
\cref{sec:Discussion} \ace{Is this the sec you are thinking of?}. For
now, we will study how to construct one-loop integrands consistent
with the color-kinematics by isolating the cubic sector of the state
sum above.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box (2box) and penta-triangle (pt) as the
Jacobi basis graphs and building a the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form:
\begin{equation}
  \mathcal{N}_{\text{2box}},\mathcal{N}_{\text{pt}}\in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  In our labeling of the two loop Jacobi-basis
diagrams, shown in \cref{eq:jacobi-basis}, we find that the power
counting bound proposed by Bern, Davies, and Nohle leads to an initial
ansatz with 10,000 terms in the double-box, and 9849 terms for the
penta-triangle.  Without power counting restrictions, both the
double-box and the penta-triangle ansatz have an ansatz with 10,010
terms each.  Imposing diagram automorphisms and maximal-cut gauge
invariance on the two basis diagrams, we reduce the number of terms to
XX for the double-box, and YY for the penta-triangle.  In
Ref.~\cite{Bern:2015ooa}, the unitarity matching constraints were
imposed using a spanning set of minimal physical cuts, shown in
\cref{fig:ym-spanning}.  Instead of also directly going to the
spanning cuts, we work within the framework of the Method of Maximal
Cuts \citepls{} in order to identify the simplest cut that is in
tension with the kinematic Jacobi relations.

\begin{figure}
  \begin{center}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \LMCut
      \end{center}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \PhysicalCutOne{}{}{}{}
      \end{center}
    \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \begin{center}
        \PhysicalCutTwo{}{}{}{}
      \end{center}
    \end{subfigure}
  \end{center}
  \caption{The three spanning physical unitarity cuts of pure
    Yang-Mills at 2 loops}
  \label{fig:ym-spanning}
\end{figure}

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut. \jm{Insert clip of Gordon Ramsay saying ``delicious'' here.}

\section{Conclusions and Outlook}\label{conclusions}
There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a 2-loop integrand for semi-abelian YM theory,
that is, the theory is 1-loop exact.  Obtaining a globally color-dual
integrand beyond 1-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual 2-loop integrands.}  \ace{I want to
  rephrase especially the last sentence.}
